Secure Composition of Robust and Optimising Compilers

### 2025-02-04 icfp/sections/intro.tex
sketch of intro

Programming languages provide security enforcements, e.g., Rust memory safety by means of ownership semantics.

Compilers may fail to preserve these enforcements, even when proven correct.

>>>>

As of today, there exist a number of programming languages that enforce security-relevant properties.
One of the most used programming languages in that space is Rust~\cite{}, which guarantees basic memory safety.
That is, there is a guarantee that all memory accesses are {\em temporally} as well as {\em spatially} safe. 
Temporal memory safety is the absence of use-after-free and use-after-reallocation bugs and necessitates tracking of pointer provenance\footnote{Due to potential use-after-reallocation bugs.}. %TODO: example?
Provenance of pointers is also crucial information, together with bounds, for spatial memory safety, which ensures that all accesses are within bounds of an allocated object.

Compilers may fail to preserve these enforcements, even when proven correct or thoroughly tested~\cite{}.
While properties can be enforced at source-level by means of static analyses, after compiling to languages without such abstractions, the properties may break.
For example, RISC-V, a possible target-language of the most commonly used Rust compiler, is untyped and provides neither memory safety nor type safety guarantees.
Intuitively, this is why we want compilers to be {\em correct}, so that they provably preserve the meaning of the source-language.
Unfortunately, software is seldolmy developed in isolation and programs may be linked with target-level libraries after compilation.
These libraries can be compiled with different compilers or may even be hand-written, so the correctness-guarantees provided by one compiler may not hold.
